---
title: "course project"
author: "m.pryidun"
date: "2023-02-08"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages:

```{r}
library(caret)
library(randomForest)
library(dplyr)
library(corrplot)
```


# Set urls:
```{r}
trainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

# Read in the data 
```{r}
download.file(trainURL, destfile = "pml-training.csv", method = "curl")
download.file(testURL, destfile = "pml-testing.csv", method = "curl")
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

# Remove na and zero values:
```{r}
NAChecker <- function(x){unlist(apply(x, 2, function(x){length(which(!is.na(x)))}))}
NDataPoints <- NAChecker(training)

CompleteVariable <- c()
for(i in 1:length(NDataPoints)){
  if(NDataPoints[[i]]==nrow(training)){
    CompleteVariable <- c(CompleteVariable, names(training)[i])
  }
}

trainingSet <- training[, names(training) %in% CompleteVariable]

nzv <- nearZeroVar(trainingSet, saveMetrics = TRUE)

myVar <- rownames(subset(nzv, nzv==FALSE))
print(myVar)
```

# Create a new data set with newly identified set of variables and remove the first 6 columns which would not be used for prediction:
```{r}
myVar <- myVar[-(1:6)]
trainingData <- select(trainingSet, one_of(myVar))
```

# Slice dataset for validationt:
```{r}
inTrain <- createDataPartition(y=trainingData$classe, p=0.6, list=FALSE)


trainingPart <- trainingData[inTrain,]
validationPart <- trainingData[-inTrain,]
```
# Check Relationahsips Among Variables:
```{r}
varCorr <- round(cor(trainingPart[sapply(trainingPart, is.numeric)]), 4)

par(ps=5)
corrplot.mixed(varCorr, order="hclust", tl.col="black", diag="n", tl.pos="lt", lower="circle", upper = "number", tl.cex=1.5, mar=c(1, 0, 1, 0))
```
# Principal Component Analysis
```{r}

reduced <- preProcess(trainingPart[,-53], method = "pca")
trainingPCA <- predict(reduced, trainingPart[,-53])
validationPCA <- predict(reduced, validationPart[,-53])
print(reduced)
```
# Build a Random Forest Model Without PCA

```{r}

modelRF2 <- train(classe ~., method="rf", data=trainingPart, trControl = trainControl(method="cv", number=4), ntree=100, importance =TRUE)

par(ps=5)
varImpPlot(modelRF2$finalModel, sort = TRUE, type = 1, pch=19, col=12, cex=1, main="Importance of Predictor Variables in Random Forest Model")
```

# Caculate the Accuracy of the Model
```{r}
modelRF2Val <- predict(modelRF2, validationPart)
modelRF2Acc <- round(postResample(validationPart$classe, modelRF2Val)[[1]], 4)
modelRF2Acc
```
#Final Test
```{r}
modelRF2Test <- predict(modelRF2, testing)
modelRF2Test
```